---
title: "Istanbul Public Transportation Network Analysis"
author: "Ahmet Emre Aladag"
date: "10 July 2015"
output: html_document
---

# Unweighted Network

## Loading Data

Let's load libraries, stop and line data.

```{r}
stops = read.csv("../crawler/output/stop.csv")
stops = unique(stops)
lines = read.csv("../crawler/output/linedetail.csv")
```


## Edge Construction

Now create edges by connecting adjacent stops in each line:

```{r}
newedges = list()
k = 1
for (i in 1:nrow(lines)) {
  line_name = as.character(lines$cdk_id[i])
  line_stops = strsplit(as.character(lines$stop_list[i]), ";")[[1]]
  for (j in 1:(length(line_stops)-1)){
    newedges[[k]] = c(line_stops[j], line_stops[j+1], line_name)
    k = k + 1
  }
}
```

Then we can convert this edge list into a matrix with columns (V1, V2, Line):

```{r}
matrix_edges = matrix(data=NA, nrow=length(newedges), ncol=3)

for (i in 1:length(newedges)){
  matrix_edges[i, 1] = newedges[[i]][1]
  matrix_edges[i, 2] = newedges[[i]][2]
  matrix_edges[i, 3] = newedges[[i]][3]
}
```

Now we convert this matrix into a data frame with columns V1 (From), V2 (To) and Line (Edge label).

```{r}
frame_edges = as.data.frame(matrix_edges, stringsAsFactors = TRUE, row.names=c("from", "to", "line"))
colnames(frame_edges) = c("V1","V2","Line")
```

## Constructing Graph

Now that we have edges and nodes, we can construct a graph object.

```{r}
library("igraph")
g = graph.data.frame(frame_edges, TRUE, stops$cdk_id)
V(g)$Latitude = stops$lat
V(g)$Longitude = stops$lon
V(g)$size = 0.5
V(g)$label = as.character(stops$name)
V(g)$labels = as.character(stops$name)
E(g)$label = as.character(frame_edges$Line)
```

## Drawing the Map

Now create an aerial map:
```{r}
library("ggmap")
location = c(mean(stops$lon), mean(stops$lat))
map = get_map(location, maptype="watercolor", source="osm", zoom=10)
```

Now draw the spatial map and plot nodes/edges on it:

```{r}
library(popgraph)
p = ggmap(map)
p = p + geom_nodeset( aes(x=Longitude, y=Latitude, size=size, label=label), g)
p = p + geom_edgeset( aes(x=Longitude, y=Latitude, shape=label, label=label), g, color="yellow", alpha=0.3)
p + xlab("Longitude") + ylab("Latitude")
# ggsave ("map.png", dpi = 200)
```

# Weighted Network

## Construct the graph

We construct the graph and add an additional weight parameter to edges. Edges will be 0-1 normalized log of the frequencies (number of buses pass through that edge).

```{r}
library(plyr)
library(ppls)
range01 <- function(x){(x-min(x))/(max(x)-min(x))}
freqs = count(frame_edges, vars=c("V1","V2"))
g2 = graph.data.frame(freqs, TRUE, stops$cdk_id)
V(g2)$Latitude = stops$lat
V(g2)$Longitude = stops$lon
V(g2)$size = 0.5
V(g2)$label = as.character(stops$name)
V(g2)$labels = as.character(stops$name)
E(g2)$freq = as.numeric(freqs$freq)
E(g2)$weight = as.numeric(range01(log(as.numeric(freqs$freq))))
E(g2)$betweenness = edge.betweenness.estimate(g2, directed=TRUE, cutoff=10, weights = E(g2)$freq)
E(g2)$logbetweenness = log(E(g2)$betweenness)
```

## Draw the map

Now we draw the map with the edge thickness and color adjusted to edge weight (bus frequency).

```{r}
location = c(mean(stops$lon), mean(stops$lat))
map11 = get_map(location, maptype="watercolor", source="osm", zoom=11)

p2 = ggmap(map11)
p2 = p2 + geom_nodeset( aes(x=Longitude, y=Latitude, size=size, label=label), g2)
p2 = p2 + geom_edgeset( aes(x=Longitude, y=Latitude, shape=label, label=label, size=weight, color=weight), g2, alpha=1.0) + scale_colour_continuous(trans = 'reverse')
p2 + xlab("Longitude") + ylab("Latitude")
#ggsave ("map2.png", dpi = 200)
```

# Edge Betweenness

Lets see the histogram of the edge betweenness values that are greater than 1:

```{r}
hist(E(g2)$logbetweenness)
```

Wee see that busiest edges are ones above logbetweenness 3. So remove less busy ones.

```{r}
small_edges = subset(E(g2), E(g2)$logbetweenness < 3)
g3 = delete.edges(g2, small_edges)
```

Now plot the map according to logarithmic edge betweenness:

```{r}
location = c(mean(stops$lon), mean(stops$lat))
map12 = get_map(location, maptype="watercolor", source="osm", zoom=12)
p3 = ggmap(map12)
p3 = p3 + geom_nodeset( aes(x=Longitude, y=Latitude, size=size, label=label), g3)
p3 = p3 + geom_edgeset( aes(x=Longitude, y=Latitude, shape=label, label=label, size=logbetweenness, color=logbetweenness), g3, alpha=1.0) + scale_colour_continuous(trans = 'reverse')
p3 + xlab("Longitude") + ylab("Latitude")
#ggsave ("map3.png", dpi = 200)
```

We can see that two bridges of Istanbul have the highest betweenness values. This means that they have very important roles in Istanbul transportation and a failure in the service would cause lots of delays. Next, we can see 
roads Besiktas-Levent, Sisli-Kagithane, around Sisli, Halic Bridge, Yenisahra E-5. Then we can see the highway (on the rightmost vertical) connecting E-5 to north bridge.


## Node Betweenness

Calculate node betweenness and sort nodes by that:

```{r}
V(g2)$betweenness = betweenness(g2, v=V(g2), directed = TRUE, weights = E(g2)$freq)
V(g2)$logbetweenness = log(V(g2)$betweenness)
dd = get.data.frame(g2, what=c("vertices"))

sorted_list = dd[order(dd$betweenness, decreasing = TRUE),]

for (i in 1:20 ) {
  print(sorted_list$label[i])
}

```

Now plot the map according to node betweenness:

```{r}
location = c(mean(stops$lon), mean(stops$lat))
map10 = get_map(location, maptype="watercolor", source="osm", zoom=10)
p4 = ggmap(map10)
p4 = p4 + geom_nodeset( aes(x=Longitude, y=Latitude, size=betweenness, label=label), g2, alpha=0.7)
p4 + xlab("Longitude") + ylab("Latitude")
#ggsave ("map4.png", dpi = 200)
```

This time adjust alpha with 0-1 normalized node betweenness to see major ones. See in a more detailed map:

```{r}
location = c(mean(stops$lon), mean(stops$lat))
map12 = get_map(location, maptype="watercolor", source="osm", zoom=12)
p5 = ggmap(map12)
p5 = p5 + geom_nodeset( aes(x=Longitude, y=Latitude, size=betweenness, label=label), g2, alpha=range01(V(g2)$betweenness))
p5 + xlab("Longitude") + ylab("Latitude")
#ggsave ("map5.png", dpi = 200)
```

